### 2020-04-10 docs/ICTT2019/RMethod_ictt26.tex
First implementation of the collision probability solver in annular geometries; ICTT26 doc update; bug fix of tracking algorithm; add a test in bare and empty bodies for escape and collision probs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{ictt26}
%
%  various packages that you may wish to activate for usage
\usepackage{graphicx}
\usepackage{tabls}
\usepackage{afterpage}
\usepackage{lastpage}
%\usepackage{cites}
%\usepackage{epsf}
\usepackage{amsmath, amssymb}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{arrows,calc}
\usepackage{subcaption}
%
% Math commands
\newcommand{\mcl}[1]{\ensuremath{\Lambda_{#1}}}
\newcommand{\cl}[1]{\ensuremath{\lambda_{#1}}}
\newcommand{\partialx}[1]{\ensuremath{\frac{\partial {#1}}{\partial x}}}
\newcommand{\derivx}[1]{\ensuremath{\frac{d {#1}}{dx}}}
\newcommand{\aflux}[2]{\ensuremath{\psi\left({#1}, {#2} \right)}}
\newcommand{\cs}[2]{\ensuremath{\sigma_{#1}\left({#2}\right)}}
\newcommand{\src}[1]{\ensuremath{q\left({#1}\right)}}
\newcommand{\csmat}[2]{\ensuremath{\sigma_{#1}^{#2}}}
\newcommand{\scatratmat}[1]{\ensuremath{c_{#1}}}
\newcommand{\sfluxnofn}{\ensuremath{\phi}}
\newcommand{\sfluxmatnofn}[1]{\ensuremath{\sfluxnofn_{#1}}}
\newcommand{\sflux}[1]{\ensuremath{\sfluxnofn \left( {#1} \right)}}
\newcommand{\sfluxmat}[2]{\ensuremath{\sfluxmatnofn{#1}\left( {#2} \right)}}
\newcommand{\sfluxrmserr}[2]{\ensuremath{\sfluxnofn_{#1}^{#2}}}
\newcommand{\ensavg}[1]{\ensuremath{\left<{#1}\right>}}
%
% Shortcut commands
\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqnolabel}[1]{(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\eqsthru}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
\newcommand{\eqsand}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
\newcommand{\tbl}[1]{Table~\ref{#1}}
\newcommand{\tblnolabel}[1]{\ref{#1}}
\newcommand{\tbls}[1]{Tables~\ref{#1}}
\newcommand{\tblsthru}[2]{Tables~\ref{#1}--\ref{#2}}
\newcommand{\tblsand}[2]{Tables~\ref{#1} and \ref{#2}}
\newcommand{\fig}[1]{Fig.~\ref{#1}}
\newcommand{\fignolabel}[1]{\ref{#1}}
\newcommand{\figs}[1]{Figs.~\ref{#1}}
\newcommand{\figsthru}[2]{Figs.~\ref{#1}--\ref{#2}}
\newcommand{\figsand}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\sxn}[1]{Section~\ref{#1}}
%
%
% Insert authors' names and short version of title in lines below
%
\newcommand{\authorHead}      % Author's names here
   {Daniele Tomatis et Al.}
\newcommand{\shortTitle}      % Short title here
   {The Ronen Method in simple 1-D problems}

@@

\begin{abstract}
In this work we apply the Ronen method to resolve the neutron transport in simple homogeneous problems. Slab, cylindrical and spherical geometries are studied. This method demands successive resolutions of the diffusion equation, where the local diffusion constants are modified in order to reproduce new estimates of the currents by a transport operator. The diffusion solver employs here finite differences and the transport-corrected currents are forced in the numerical scheme by means of drift terms, like in the CMFD technique. The solution is compared against reference results provided by the collision probability method. Boundary conditions are discussed introducing proper approximations to save the particle balance. Slow convergence is reported on the scalar flux, although the results match the reference solutions in the limit of fine meshes and far from the bare boundary.
%
%In this work we apply the Ronen method in the slab geometry filled with heterogeneous media. Neutron diffusion is solved by a nodal method on a coarse mesh, while reference solutions of the neutron distribution are provided by a discrete ordinate transport solver. The CMFD technique is used for both the implementation of the nodal solver and for the corrections of the diffusion coefficient given by the Ronen method. The presence of discontinuity factors for equivalencing the diffusion solution is investigated jointly with the Ronen method. Results are presented for different levels of material heterogeneity at the slab interfaces.
\end{abstract}
%
% SET RAGGED RIGHT MARGIN
%
\raggedright

@@

%\section{NODAL DIFFUSION}
%\label{sec:nd}
%
%Daniele: this topic is postponed to future works.
%
\section{CURRENT ESTIMATION}
\label{sec:ce}

@@

Multiplication of Eq. \ref{eq:NTE} by $\mu$ and the integration over $[-1, 1]$ provides the expression for the current, see \cite{tomatis2011application}. The integration on $\mu$ yields integral exponential functions by a simple change of variable ($\mu = \pm 1/u$, according to the sign of $\mu$) \cite{AS1964handbook}:
\begin{equation*}
E_n(\tau) = \int_0^1{d\mu\, e^{-\tau / \mu} \mu^{n-2}} = \int_1^\infty{du\, e^{-\tau u}u^{-n}}, \: n \geq 0.
\end{equation*}
These special functions are then used to compute the current at the cell interfaces for all $i$:
\begin{equation}
\begin{split}
% J_g(x_{i+1/2}) &=
J_{g,i+1/2} &= \int_0^{ 1}{d\mu\, \mu \varphi_g(a, \mu) e^{-\tau_g(a, x_{i+1/2})/\mu}}
             - \int_0^{-1}{d\mu\, \mu \varphi_g(b, \mu) e^{-\tau_g(b, x_{i+1/2})/\mu}}\\
  & + \sum_{j=0}^{I-1} \left[  \sign(x_{i+1/2} - x_{j+1/2}) \frac{q_{0,g,j}}{2} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{2}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } \right.\\
  & \left. \qquad+ \frac{3}{2}q_{1,g,j} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{3}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) }\right],
\end{split}
\label{eq:J_tr}
\end{equation}
where $q_{(.),g,j}$ are the moments of the souce $q_g$ from Eq. \ref{eq:srcg} averaged in the volume of the $j$-th cell:
\begin{equation*}
\label{eq:srclg}
\begin{split}
   q_{0,g,j} &= \sum_{g'}\left(\sigma_{s,0,g' \rightarrow g,j} +
                              \chi_g \nu\sigma_{f,g',j}
                        \right) \phi_{g',j},\\
   q_{1,g,j} &= \sum_{g'}\sigma_{s,1,g' \rightarrow g,j}J_{g',j}.
\end{split}
\end{equation*}
The current appearing in $q_{1,g,j}$, that is in case of linearly anisotropic scattering, is also to be considered as volume-averaged in the cell $j$. The optical lengths show the subscript $g$ because they are evaluated with the corresponding total cross section $\sigma_{t,g}$. For the property $E'_{n+1}(\ell) = -E_n(\ell)$, the spatial integrals of the integral exponential functions can be solved analytically by integrating on $\tau$:
\begin{equation*}
\begin{aligned}
\int_{x_{j-1/2}}^{x_{j+1/2}}{ dx' E_{n}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } =
\frac{\sign(x_{i+1/2} - x_{j+1/2})}{\sigma_{t,g,j}} \cdot \\
\left[
  E_{n+1} \left(\lvert \tau_g(x_{j+1/2}, x_{i+1/2}) \rvert \right)
 -E_{n+1} \left(\lvert \tau_g(x_{j-1/2}, x_{i+1/2}) \rvert \right)
\right].
\end{aligned}
\end{equation*}
These quantities can be computed for $i > j$ only, because the others are simply opposite in sign.

{\color{red}Roy-Erez: Please add a paragraph or reformulate the previous one to introduce the escape probabilities in the slab.}

@@

As well, integration of Eq. \ref{eq:ITE_xy} over the solid angle $d\mathbf{\Omega} = \sin \theta d\omega d \theta$ yields the scalar flux. The theory of collision probability methods originates from the use of isotropic sources, which are also considered as uniform in the cells of the spatial mesh. Two integrations in space arise for each direction $\mathbf{\Omega}$, connecting the source in region $j$ to the flux (or its total reaction rate) in region $i$ by means of its collision probability. The incoming flux is considered as isotropic too, or linearly anisotropic in angle as in \cite{hebert2009applied} to relate the partial entering and outgoing currents to the only first flux moments. These partial currents on the boundary must satisfy a condition of albedo. Transfer probabilities across this outer surface are then derived to take into account of incoming particles in the equation for the $i$-th scalar flux. Escape probabilities accounts for neutrons produced in every region and leaving the outer surface uncollided.

The projections of partial currents on the outward normal $\hat{n}$ to a given surface are obtained from integration of Eq. \ref{eq:ITE_xy} with the weight $| \hat{n} \cdot \mathbf{\Omega} | = \sin \theta | \hat{n} \cdot \mathbf{\Omega}_p |$ on $\hat{n} \cdot \mathbf{\Omega} \lessgtr 0$. Higher moments are obtained by the weights associated to the corresponding spherical harmonics. Integrations along the polar angle are resolved with the introduction of the Bickley-Naylor functions \cite{amos1983uniform}:
\begin{equation}
\text{Ki}_n(\tau) = \int_0^{\pi/2}{ d\theta \sin^{n-1} \theta \exp\left(-\frac{\tau}{\sin \theta}\right) },\: n \geq 0.
\label{eq:Kin}
\end{equation}
The properties used for the computation of these functions are available elsewhere \cite{lewis1984computational,hebert2009applied,amos1983algorithm}.

\begin{subequations}
After introducing these arguments for a bare cylinder, the current leaving the cylindrical surface at $\mathbf{r}$ (with direction $\hat{n}$) given by neutrons flying along $\mathbf{\Omega}_p$ is:
\begin{equation}
J^+(\mathbf{r}) = \frac{1}{2\pi}\int_{W^+}{
  d\omega\, | \hat{n} \cdot \mathbf{\Omega}_p | \,
  \int_0^S ds\, \text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r} - s \mathbf{\Omega}_p \right)\right]
  q(\mathbf{r} - s \mathbf{\Omega}_p)},
\label{eq:currp_cylinder1}
\end{equation}
with $W^+ = \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p > 0\right\}$. Using the surface element $dA' = s d\omega ds$ and $\mathbf{r}'_p = \mathbf{r} - s \mathbf{\Omega}_p$, it is also:
\begin{equation}
J^+(\mathbf{r}) = \int_{W^+ \times [0, S]}{
  dA'\, \hat{n} \cdot (\mathbf{r} - \mathbf{r}'_p) \,
  \frac{\text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r}'_p \right)\right]}{2 \pi \lvert \mathbf{r} - \mathbf{r}'_p \rvert^2}
  q(\mathbf{r}'_p)}.
\label{eq:currp_cylinder2}
\end{equation}
\label{eq:currp_cylinder}
\end{subequations}
$J^-$ is obtained with integration over $w \in W^- =  \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p < 0\right\}$ instead.

Another integration on the total surface $S(r) = 2\pi r$ is necessary to obtain the current in the 1-D frame. For a given angle $\omega$, we have then volume integrals to solve numerically along many parallel lines called tracks, see Fig. \ref{fig:cyltracks}. Furthermore, the tracks are identical for any angle $\omega$ in the 1-D cylindrical geometry simplifying the integrals in Eqs. \ref{eq:currp_cylinder}, which can also be limited to half portion of the cylinder thanks to its symmetry. Finally, the outgoing and incoming currents at $r$ become:
\begin{equation}
J^\pm(r) = \pm \frac{1}{\pi r^2} \int_0^r { dh\, y \int_{-Y}^{\pm y}{
  d \ell \, \text{Ki}_2\left[\tau( y, y - \ell )\right] q(h, y - \ell)}}
\label{eq:currp_cyl_atr}
\end{equation}
with $y(r,h) =\sqrt{r^2 - h^2}$, $Y = y(R,h)$ and the outer radius $R$. After multiplication by $S(r)$ and still with uniform sources, Eq. \ref{eq:currp_cyl_atr} can be written in terms of the escape probability $e_{j}(r)$ for a neutron to be emitted isotropically in ring $j$ and to leave uncollided the semicylinder surface at $r$, being crossed with angle $\omega \in W^\pm$:
\begin{equation}
J^\pm(r)S(r) = \sum_j q_j V_j e^\pm_j(r) \mbox{ with }
e^\pm_j(r) = \pm \frac{2}{r V_j} \int_0^r { dh\, y \int_{L^\pm_j}{
  d \ell \, \text{Ki}_2\left[\tau( y, y - \ell )\right]}},
\label{eq:esc_prob_cyl}
\end{equation}
$L^\pm_j(h) = [-Y,\pm y] \cap V_j$ and where $V_j$ is the ring area. The integral on the track $\ell$ is written differently for the convex and for the concave parts of the rings with respect to the direction of flight, see for instance track 1 and 2 in Fig. \ref{fig:cyltracks} where the same notation of Fig. \ref{fig:mesh1D} is used for the radial mesh. The current leaving the $i$-th ring at $r_{i+1/2}$ gets all source contributions weighted with the quantities $\varepsilon^+_{i+1/2,j} = V_j e^+_{i+1/2,j}$:
%
\newlength{\plussign}
\settowidth{\plussign}{$+$}
%
\begin{equation}
\label{eq:esc_prob_cyij}
  \begin{gathered}
  \varepsilon^+_{i+1/2,j} = \frac{2}{r_{i+1/2}} \cdot\\
  %\left\{
  \begin{cases}
  \displaystyle ~\int_0^{r_{i-1/2}} { dh\, y
    \int_0^{\ell_j}{ d\ell \bigg[ }}\\[3mm] \qquad
      \displaystyle \hspace{\the\plussign}
        \text{Ki}_2\left(2 \sigma_i \ell_i + \tau_{ii} + \tau_{ij} + \sigma_j \ell \right)
        \text{ if } i < j \\[3mm] \qquad
      \displaystyle +
        \text{Ki}_2\left(\sigma_i \ell_i + (1 - \delta_{ij})(\tau_{ij} + \sigma_j \ell_j) + \tau_{jj} + \sigma_j \ell \right)
        \text{ if } i \geq j \\[3mm] \qquad
      \displaystyle +
        \text{Ki}_2\left((1 - \delta_{ij})(\sigma_i \ell_i + \tau_{ij}) + \sigma_j \ell \right)
        \text{ if } i \geq j \\
%
%  \displaystyle + \int_0^{r_{i-1/2}} { dh\, y
%    \int_0^{\ell_j}{ d\ell \, \left[
%      \text{Ki}_2\left(\sigma_i \ell_i + \tau_{ij} + \sigma_j \ell_j + \tau_{jj} + \sigma_j \ell \right)
%      \right. }} \\ {{ \left.
%      + \text{Ki}_2\left(\sigma_i \ell_i + \tau_{ij} + \sigma_j \ell \right) \right]} }
%  \text{ if } i \geq j\\
%
  \displaystyle \: \bigg] +\int_{r_{i-1/2}}^{r_{i+1/2}} { dh\, y
    \int_0^{\ell_j}{ d\ell \, \text{Ki}_2\left((1-\delta_{ij})(\sigma_i \ell_i + \tau_{ij}) + \sigma_j \ell \right)} }
  \text{ if } i \leq j,\\
  \end{cases}
  %\right.
  \end{gathered}
\end{equation}
using the Kronecker function $\delta_{ij} = 1$ only if $i=j$ and null otherwise. $\tau_{ij}$ and $\tau_{ii}$ are respectively the optical lengths between the rings $i$ and $j$, and across the outer diameter of ring $i$. Reciprocity and conservation properties can be applied only after calculating the collision probabilities. The integrals involving the Ki function are solved thanks to the property $d_{\tau}\text{Ki}_n(\tau) = -\text{Ki}_{n-1}(\tau)$.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{ictt26}
%
%  various packages that you may wish to activate for usage
\usepackage{graphicx}
\usepackage{tabls}
\usepackage{afterpage}
\usepackage{lastpage}
%\usepackage{cites}
%\usepackage{epsf}
\usepackage{amsmath, amssymb, bm}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{arrows,calc}
\usepackage{subcaption}
%
% Math commands
\newcommand{\mcl}[1]{\ensuremath{\Lambda_{#1}}}
\newcommand{\cl}[1]{\ensuremath{\lambda_{#1}}}
\newcommand{\partialx}[1]{\ensuremath{\frac{\partial {#1}}{\partial x}}}
\newcommand{\derivx}[1]{\ensuremath{\frac{d {#1}}{dx}}}
\newcommand{\aflux}[2]{\ensuremath{\psi\left({#1}, {#2} \right)}}
\newcommand{\cs}[2]{\ensuremath{\sigma_{#1}\left({#2}\right)}}
\newcommand{\src}[1]{\ensuremath{q\left({#1}\right)}}
\newcommand{\csmat}[2]{\ensuremath{\sigma_{#1}^{#2}}}
\newcommand{\scatratmat}[1]{\ensuremath{c_{#1}}}
\newcommand{\sfluxnofn}{\ensuremath{\phi}}
\newcommand{\sfluxmatnofn}[1]{\ensuremath{\sfluxnofn_{#1}}}
\newcommand{\sflux}[1]{\ensuremath{\sfluxnofn \left( {#1} \right)}}
\newcommand{\sfluxmat}[2]{\ensuremath{\sfluxmatnofn{#1}\left( {#2} \right)}}
\newcommand{\sfluxrmserr}[2]{\ensuremath{\sfluxnofn_{#1}^{#2}}}
\newcommand{\ensavg}[1]{\ensuremath{\left<{#1}\right>}}
%
% Shortcut commands
\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqnolabel}[1]{(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\eqsthru}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
\newcommand{\eqsand}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
\newcommand{\tbl}[1]{Table~\ref{#1}}
\newcommand{\tblnolabel}[1]{\ref{#1}}
\newcommand{\tbls}[1]{Tables~\ref{#1}}
\newcommand{\tblsthru}[2]{Tables~\ref{#1}--\ref{#2}}
\newcommand{\tblsand}[2]{Tables~\ref{#1} and \ref{#2}}
\newcommand{\fig}[1]{Fig.~\ref{#1}}
\newcommand{\fignolabel}[1]{\ref{#1}}
\newcommand{\figs}[1]{Figs.~\ref{#1}}
\newcommand{\figsthru}[2]{Figs.~\ref{#1}--\ref{#2}}
\newcommand{\figsand}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\sxn}[1]{Section~\ref{#1}}
%
%
% Insert authors' names and short version of title in lines below
%
\newcommand{\authorHead}      % Author's names here
   {Daniele Tomatis et Al.}
\newcommand{\shortTitle}      % Short title here
   {The Ronen Method in simple 1-D problems}

@@

\begin{abstract}
In this work we apply the Ronen method to resolve the neutron transport in simple homogeneous problems. Slab, cylindrical and spherical geometries are studied. This method demands successive resolutions of the diffusion equation, where the local diffusion constants are modified in order to reproduce new estimates of the currents by a transport operator. The diffusion solver employs here finite differences and the transport-corrected currents are forced in the numerical scheme by means of drift terms, like in the CMFD scheme. Boundary conditions are discussed introducing proper approximations to save the particle balance in case of reflection in the slab. The solution from the Ronen iterations is compared against reference results provided by the collision probability method (CPM). More accurate estimates of the currents are provided by integral equations using first flight escape probabilities. Slow convergence on the scalar flux is analyzed, although the results match the reference solutions in the limit of fine meshes and far from the bare boundary.
%
%In this work we apply the Ronen method in the slab geometry filled with heterogeneous media. Neutron diffusion is solved by a nodal method on a coarse mesh, while reference solutions of the neutron distribution are provided by a discrete ordinate transport solver. The CMFD technique is used for both the implementation of the nodal solver and for the corrections of the diffusion coefficient given by the Ronen method. The presence of discontinuity factors for equivalencing the diffusion solution is investigated jointly with the Ronen method. Results are presented for different levels of material heterogeneity at the slab interfaces.
\end{abstract}
%
\clearpage
\tableofcontents
\clearpage
%
% SET RAGGED RIGHT MARGIN
%
\raggedright

@@

\section{CURRENT ESTIMATION}
\label{sec:ce}

@@

Multiplication of Eq. \ref{eq:NTE} by $\mu$ and the integration over $[-1, 1]$ provides the expression for the current, see \cite{tomatis2011application}. The integration on $\mu$ yields integral exponential functions by a simple change of variable ($\mu = \pm 1/u$, according to the sign of $\mu$) \cite{AS1964handbook}:
\begin{equation*}
E_n(\tau) = \int_0^1{d\mu\, e^{-\tau / \mu} \mu^{n-2}} = \int_1^\infty{du\, e^{-\tau u}u^{-n}}, \: n \geq 0.
\end{equation*}
These special functions are then used to compute the current at the cell interfaces for all $i$:
\begin{equation}
\begin{split}
% J_g(x_{i+1/2}) &=
J_{g,i+1/2} &= \int_0^{ 1}{d\mu\, \mu \varphi_g(a, \mu) e^{-\tau_g(a, x_{i+1/2})/\mu}}
             - \int_0^{-1}{d\mu\, \mu \varphi_g(b, \mu) e^{-\tau_g(b, x_{i+1/2})/\mu}}\\
  & + \sum_{j=0}^{I-1} \left[  \sign(x_{i+1/2} - x_{j+1/2}) \frac{q_{0,g,j}}{2} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{2}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } \right.\\
  & \left. \qquad+ \frac{3}{2}q_{1,g,j} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{3}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) }\right],
\end{split}
\label{eq:J_tr}
\end{equation}
where $q_{(.),g,j}$ are the moments of the souce $q_g$ from Eq. \ref{eq:srcg} averaged in the volume of the $j$-th cell:
\begin{subequations}
\label{eq:srclg}
\begin{align}
   q_{0,g,j} &= \sum_{g'}\left(\sigma_{s,0,g' \rightarrow g,j} +
                              \chi_g \nu\sigma_{f,g',j}
                        \right) \phi_{g',j},\label{eq:srclg0}\\
   q_{1,g,j} &= \sum_{g'}\sigma_{s,1,g' \rightarrow g,j}J_{g',j}.\label{eq:srclg1}
\end{align}
\end{subequations}
The current appearing in $q_{1,g,j}$, that is in case of linearly anisotropic scattering, is also to be considered as volume-averaged in the cell $j$. The optical lengths show the subscript $g$ because they are evaluated with the corresponding total cross section $\sigma_{t,g}$. For the property $E'_{n+1}(\ell) = -E_n(\ell)$, the spatial integrals of the integral exponential functions can be solved analytically by integrating on $\tau$:
\begin{equation*}
\begin{aligned}
\int_{x_{j-1/2}}^{x_{j+1/2}}{ dx' E_{n}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } =
\frac{\sign(x_{i+1/2} - x_{j+1/2})}{\sigma_{t,g,j}} \cdot \\
\left[
  E_{n+1} \left(\lvert \tau_g(x_{j+1/2}, x_{i+1/2}) \rvert \right)
 -E_{n+1} \left(\lvert \tau_g(x_{j-1/2}, x_{i+1/2}) \rvert \right)
\right].
\end{aligned}
\end{equation*}
These quantities can be computed for $i > j$ only, because the others are simply opposite in sign (anti-symmetric). Eq. \ref{eq:J_tr} can be reformulated for the partial currents, thus considering only the contributions coming from the different sides of $x_{i+1/2}$. Omitting the index on groups and keeping only isotropic sources, this is:
\begin{equation}
\label{eq:pcslab}
  J^\pm_{i+1/2} = \sum_{j=1}^I{q_{0,j} \Delta_i \tilde{e}^\pm_{i+1/2,j} + J^\pm(x_b) \tilde{t}_{i+1/2, x_b}}.
\end{equation}
$\tilde{e}_{i+1/2, j}$ represents the probability of a neutron emitted isotropically in $\Delta_i$ to escape uncollided at $x_{i+1/2}$, along its positive or negative direction of flight. $\tilde{t}_{i+1/2, x_b}$ is the transmission probability of a neutron to enter isotropically at $x_b$ and to cross the interface at $x_{i+1/2}$ without colliding hitherto.

@@

As well, integration of Eq. \ref{eq:ITE_xy} over the solid angle $d\mathbf{\Omega} = \sin \theta d\omega d \theta$ yields the scalar flux at the point $\mathbf{r}$. The theory of collision probability methods originates from the use of isotropic sources, which are also considered as uniform in the cells of the spatial mesh. The use of anisotropic sources with spatial variation in the same cells, for instance polynomial-like, is possible but leads to much more difficult expressions to solve. Two integrations in space arise for each direction $\mathbf{\Omega}$, connecting the source in region $j$ to the flux (or its total reaction rate) in region $i$ by means of its collision probability. The incoming flux is considered as isotropic too, or linearly anisotropic in angle as in \cite{hebert2009applied} to relate the partial entering and outgoing currents to the only first flux moments. These partial currents on the boundary must satisfy a general condition of albedo.

Using standard nomenclature of collision probability methods, transfer probabilities are derived to take into account of the contributions in the equation for the $i$-th scalar flux (or total rate) of neutrons leaving a given surface. Escape probabilities account for neutrons produced in a given region and leaving a surface uncollided instead.

The projections of partial currents on the outward normal $\hat{n}$ to a given surface are obtained from integration of Eq. \ref{eq:ITE_xy} in $d\mathbf{\Omega}$ with the weight $| \hat{n} \cdot \mathbf{\Omega} | = \sin \theta | \hat{n} \cdot \mathbf{\Omega}_p |$ on $\hat{n} \cdot \mathbf{\Omega} \lessgtr 0$. Higher moments can be obtained by the weights associated to the corresponding spherical harmonics. Integrations along the polar angle are resolved with the introduction of the Bickley-Naylor functions \cite{amos1983uniform}:
\begin{equation}
\text{Ki}_n(\tau) = \int_0^{\pi/2}{ d\theta \sin^{n-1} \theta \exp\left(-\frac{\tau}{\sin \theta}\right) },\: n \geq 0.
\label{eq:Kin}
\end{equation}
The properties used for the computation of these functions are available elsewhere \cite{lewis1984computational,hebert2009applied,amos1983algorithm}.

\begin{subequations}
After introducing these arguments for a bare cylinder, the current leaving the cylindrical surface at $\mathbf{r}$ (with direction $\hat{n}$) given by neutrons flying along $\mathbf{\Omega}_p$ is:
\begin{equation}
J^+(\mathbf{r}) = \frac{1}{2\pi}\int_{W^+}{
  d\omega\, | \hat{n} \cdot \mathbf{\Omega}_p | \,
  \int_0^S ds\, \text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r} - s \mathbf{\Omega}_p \right)\right]
  q(\mathbf{r} - s \mathbf{\Omega}_p)},
\label{eq:currp_cylinder1}
\end{equation}
with $W^+ = \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p > 0\right\}$. Using the surface element $dA' = s d\omega ds$ and $\mathbf{r}'_p = \mathbf{r} - s \mathbf{\Omega}_p$, it is also:
\begin{equation}
J^+(\mathbf{r}) = \int_{W^+ \times [0, S]}{
  dA'\, \hat{n} \cdot (\mathbf{r} - \mathbf{r}'_p) \,
  \frac{\text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r}'_p \right)\right]}{2 \pi \lvert \mathbf{r} - \mathbf{r}'_p \rvert^2}
  q(\mathbf{r}'_p)}.
\label{eq:currp_cylinder2}
\end{equation}
\label{eq:currp_cylinder}
\end{subequations}
$J^-$ is obtained with integration over $w \in W^- =  \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p < 0\right\}$ instead.% The coefficient $1/2\pi$ comes from the probability of isotropic angular redistribution applied to the source $q$.

Another integration on the total surface $S(r) = 2\pi r$ is necessary to obtain the current density in the 1-D frame. For a given angle $\omega$, we have then volume integrals to solve numerically along many parallel lines called tracks, see Fig. \ref{fig:cyltracks}. Furthermore, the tracks are identical for any angle $\omega$ in the 1-D cylindrical geometry simplifying the integrals in Eqs. \ref{eq:currp_cylinder}, which can also be limited to half portion of the cylinder thanks to its symmetry. Finally, the outgoing and incoming currents at $r$ become:
J^\pm(r) = \pm \frac{1}{\pi r^2} \int_0^r { dh\, y \int_{-Y}^{\pm y}{
  d \ell \, \text{Ki}_2\left[\tau( y, y - \ell )\right] q(h, y - \ell)}}
\label{eq:currp_cyl_atr}
with $y(r,h) =\sqrt{r^2 - h^2}$, $Y = y(R,h)$ and the outer radius $R$. After multiplication by $S(r)$ and still with uniform sources, Eq. \ref{eq:currp_cyl_atr} can be written in terms of the escape probability $e_{j}(r)$ for a neutron to be emitted isotropically in ring $j$ and to leave uncollided along the normal direction the semicylinder surface at $r$, being crossed with angle $\omega \in W^\pm$:
J^\pm(r)S(r) = \sum_j q_j V_j \left[ \pm e^\pm_j(r) \right] \mbox{ with }
e^\pm_j(r) = \frac{2}{r V_j} \int_0^r { dh\, y \int_{L^\pm_j}{
  d \ell \, \text{Ki}_2\left[\tau( y, y - \ell )\right]}},
\label{eq:esc_prob_cyl}
$L^\pm_j(h) = [-Y,\pm y] \cap V_j$ and where $V_j$ is the ring area. $\text{Ki}_2\left[\tau( y, y - \ell )\right]$ is the transport kernel expressing the probability of a neutron to travel uncollided along $\omega$ for a lenght $\tau$, given in mean free path units, from $(y - \ell)$ to $y$ \cite{stamm1983methods}. The integral on the track $\ell$ is written differently for the convex and for the concave parts of the rings with respect to the direction of flight, see for instance track 1 and 2 in Fig. \ref{fig:cyltracks} where the same notation of Fig. \ref{fig:mesh1D} is used for the radial mesh. The current leaving the $i$-th ring at $r_{i+1/2}$ gets all source contributions weighted with the quantities $\varepsilon^+_{i+1/2,j} = V_j e^+_{i+1/2,j}$:
\newlength{\plussign}
\settowidth{\plussign}{$+$}
\label{eq:esc_prob_cyij}
  \begin{gathered}
  \varepsilon^+_{i+1/2,j} = \frac{2}{r_{i+1/2}} \cdot\\
  %\left\{
  \begin{cases}
  \displaystyle ~\int_0^{r_{i-1/2}} { dh\, y
    \int_0^{\ell_j}{ d\ell \bigg[ }}\\[3mm] \qquad
      \displaystyle \hspace{\the\plussign}
        \text{Ki}_2\left(2 \sigma_i \ell_i + \tau_{ii} + \tau_{ij} + \sigma_j \ell \right)
        \text{ if } i < j \\[3mm] \qquad
      \displaystyle +
        \text{Ki}_2\left(\sigma_i \ell_i + (1 - \delta_{ij})(\tau_{ij} + \sigma_j \ell_j) + \tau_{jj} + \sigma_j \ell \right)
        \text{ if } i \geq j \\[3mm] \qquad
      \displaystyle +
        \text{Ki}_2\left((1 - \delta_{ij})(\sigma_i \ell_i + \tau_{ij}) + \sigma_j \ell \right)
        \text{ if } i \geq j \\
%  \displaystyle + \int_0^{r_{i-1/2}} { dh\, y
%    \int_0^{\ell_j}{ d\ell \, \left[
%      \text{Ki}_2\left(\sigma_i \ell_i + \tau_{ij} + \sigma_j \ell_j + \tau_{jj} + \sigma_j \ell \right)
%      \right. }} \\ {{ \left.
%      + \text{Ki}_2\left(\sigma_i \ell_i + \tau_{ij} + \sigma_j \ell \right) \right]} }
%  \text{ if } i \geq j\\
  \displaystyle \: \bigg] +\int_{r_{i-1/2}}^{r_{i+1/2}} { dh\, y
    \int_0^{\ell_j}{ d\ell \, \text{Ki}_2\left((1-\delta_{ij})(\sigma_i \ell_i + \tau_{ij}) + \sigma_j \ell \right)} }
  \text{ if } i \leq j,\\
  \end{cases}
  %\right.
  \end{gathered}
using the Kronecker function $\delta_{ij} = 1$ only if $i=j$ and null otherwise. $\tau_{ij}$ and $\tau_{ii}$ are respectively the optical lengths between the rings $i$ and $j$, and across the outer diameter of ring $i$. The radial integral includes all rings within $r_{i+1/2}$. Reciprocity and conservation properties can be applied only after calculating the collision probabilities. The integrals involving the Ki function are solved thanks to the property $d_{\tau}\text{Ki}_n(\tau) = -\text{Ki}_{n-1}(\tau)$.

@@

\appendix
\section{APPENDIX. Gauss-quadratures}
\label{sec:apx_Flurig}
